\documentclass[3p,sort&compress,11pt,fleqn,review]{elsarticle}
\bibliographystyle{elsarticle-num-names}
\usepackage{amsmath,amsfonts,amssymb,siunitx,float,subcaption,setspace,booktabs,bm,mathtools,url,algorithm,algpseudocode,lineno,lmodern}
\usepackage[fixed]{fontawesome5}
%\usepackage{easyReview}
\journal{}
\newcommand{\balpha}{\mb{\alpha}}
\newcommand{\bbeta}{\mb{\beta}}
\newcommand{\bc}{\mb{c}}
\newcommand{\bd}{\mb{d}}
\newcommand{\be}{\mb{e}}
\newcommand{\beeta}{\mb{\eta}}
\newcommand{\bgamma}{\mb{\gamma}}
\newcommand{\bn}{\mb{n}}
\newcommand{\bq}{\mb{q}}
\newcommand{\bs}{\mb{s}}
\newcommand{\bsigma}{\mb{\sigma}}
\newcommand{\bv}{\mb{v}}
\newcommand{\bvarepsilon}{\mb{\varepsilon}}
\newcommand{\bzeta}{\mb{\zeta}}
\newcommand*{\algoref}[1]{Algorithm~\ref{#1}}
\newcommand*{\ddfrac}[2]{\dfrac{\md{#1}}{\md{#2}}}
\newcommand*{\dev}[1]{\text{dev}\left(#1\right)}
\newcommand*{\diag}[1]{\text{diag}\left(#1\right)}
\newcommand*{\eqsref}[1]{Eq.~(\ref{#1})}
\newcommand*{\figref}[1]{Fig.~\ref{#1}}
\newcommand*{\mb}[1]{\bm{#1}}
\newcommand*{\md}[1]{\mathrm{d}#1}
\newcommand*{\mdd}[1]{\mathrm{d}^2#1}
\newcommand*{\mT}{\mathrm{T}}
\newcommand*{\pdfrac}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
\newcommand*{\secref}[1]{\S~\ref{#1}}
\newcommand*{\sign}[1]{\text{sign}\left(#1\right)}
\newcommand*{\tabref}[1]{Table~\ref{#1}}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}
\DeclarePairedDelimiter\norm{\lVert}{\rVert}
\makeatletter
\let\oldabs\abs\def\abs{\@ifstar{\oldabs}{\oldabs*}}
\let\oldnorm\norm\def\norm{\@ifstar{\oldnorm}{\oldnorm*}}
\newenvironment{breakablealgorithm}{\begin{center}\refstepcounter{algorithm}\hrule height.8pt depth0pt \kern2pt
\renewcommand{\caption}[2][\relax]{
{\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}
\ifx\relax##1\relax
\addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}
\else
\addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}
\fi
\kern2pt\hrule\kern2pt}}{\kern2pt\hrule\relax\end{center}}
\makeatother
\begin{document}
\linenumbers
\begin{abstract}
    \begin{linenumbers}
        In this work, we revisit the subloading surface model, as well as its extended version.
        By investigating its original formulation, we provide an alternative interpretation.
        With which, it is shown that the corresponding evolution rules of historic variables can be greatly simplified.
        The proposed evolution rules are analytically more concise and possess straightforward implications that can be easily illustrated while remain mathematically equivalent to the original ones.
        Based on the new proposal, an implicit numerical algorithm is presented.
        Compared to the previous implementation, the new algorithm is more efficient and robust.
        Additional insights regarding the unconventional (un)loading criterion and the specific form of some functions are also discussed.
    \end{linenumbers}
\end{abstract}
\begin{keyword}
    constitutive modelling\sep
    plasticity\sep
    subloading surface model\sep
    ratcheting\sep
    cyclic response
\end{keyword}
\begin{frontmatter}
    \title{On the Subloading Surface Model for Metals: Some Insights and an Efficient Numerical Implementation}
    \author[add1]{Theodore~L.~Chang\corref{tlc}}\ead{tlcfem@gmail.com}
    \cortext[tlc]{corresponding author}
    \address[add1]{IRIS Adlershof, Humboldt-Universität zu Berlin, Berlin, Germany, 12489.}
\end{frontmatter}
\input{SEC/INTRO}
\input{SEC/THEORY}
\input{SEC/EXAMPLE}
\input{SEC/CONCLUSION}
\bibliography{BIB}
\end{document}